%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
%
% Original author:
% David Contento 
%
% Important note:
% This template requires the resume.cls file to be in the same directory as the
% .tex file. The resume.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------s------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{resume} % Use the custom resume.cls style

\usepackage[left=0.65in,top=0.4in,right=0.65in,bottom=0.4in]{geometry} % Document margins
\newcommand{\tab}[1]{\hspace{.2667\textwidth}\rlap{#1}}
\newcommand{\itab}[1]{\hspace{0em}\rlap{#1}}
\name{David Contento} % Your name
\address{(+1)954-562-3893 \\ DavidFContento@gmail.com}
\address{Github.com/DContento/Past-Projects.git} 
\address{884-A 9th Avenue, Honolulu, HI 96816} % Your address

\begin{document}
\vspace{-1.3em}
%----------------------------------------------------------------------------------------
%	PROFILE SECTION
%----------------------------------------------------------------------------------------
\begin{rSection}{Profile}

With experience as an analyst in consumer banking, I am eager to put my skills to work by helping companies find creative solutions to their problems and needs. I am a self motivated, and analytically driven individual that possesses a strong understanding of data analysis, database administration, process improvement, and programming. 

\end{rSection}

%----------------------------------------------------------------------------------------
%	EDUCATION SECTION
%----------------------------------------------------------------------------------------
\begin{rSection}{Education}

{\bf Florida State University, Tallahassee, FL} \hfill {\em June 2015 - May 2018} 
\\ Bachelor of Science, Economics \hfill { \bf GPA: 3.82}
\\ Minor, General Business \hfill {\em Magna Cum Laude}
\\ 
\\{\bf University of California-Los Angeles, Los Angeles, CA} \hfill {\em July 2018 - June 2019} 
\\ Masters of Applied Economics\hfill { \bf GPA: 3.77 }

\end{rSection}

%----------------------------------------------------------------------------------------
%	Professional Experience SECTION
%----------------------------------------------------------------------------------------
\begin{rSection}{Professional Experience}

\begin{rSubsection}{First Hawaiian Bank, Honolulu, HI}{November 2021 - Present}{Risk Data  Analyst}{}
\item Assisted with Core lending system conversion and helped ensure data integrity for audit and reporting needs across all source/origination systems
\item Used Gitlab/bash along with Jira integration to track/assign issues, maintain code, and communicate with stakeholders on their respective projects. 
\item Work closely with ETL developers to make updates to SQL scripts, procedures, and views in order to maintain accurate reporting data across multiple data warehouses.
\end{rSubsection}

\begin{rSubsection}{American Savings Bank, Honolulu, HI}{July 2019 - Present}{Operational Excellence Analyst}{September 2020 - November 2021}
\item Took a consulting style approach to process improvement across the bank using a lean six sigma model to help improve efficiency to offset decreasing revenue during a time of shrinking net-interest margin.
\item Proactively coordinate with other teams in order to accomplish larger enterprise level projects such as the bank wide transition to SalesForce using an Agile development methodology.
\item Created tables/views for reporting and workflow automation using SQL, Denodo, and Cognos as well as performance tuning for all databases.
\end{rSubsection}

\vspace{-1.3em}

\begin{rSubsection}{}{}{Sales Data Analyst}{July 2019 - September 2020}
\item Generate ad hoc and scheduled reports for mid-level managers and directors of our three key business units (Investment Services, Home Loans, and Business Banking) regarding our overall retail branch performance.
\item Responsible for managing, cleaning, and gathering data from our EDW using MSSQL and administrating our Microsoft CRM platform. Also helped with data management for PCLender, Signature, and other Fiserv applications.
\item Used R, Python, and Tableau to transform, clean, and visual data as well as generate models and derive actionable analytic insight regarding our market base and retail branch performance.
\end{rSubsection}

\end{rSection}
%--------------------------------------------------------------------------------
%    Projects And Seminars
%-----------------------------------------------------------------------------------------------
\begin{rSection}{Projects}

{\bf Labor Market Efficiency in Major League Baseball}\\
Analyzed labor markets in Major League Baseball to determine systemic inefficiencies. Used regression, VAR, and other models to answer questions regarding player salary, valued offensive statistics, and the payroll effects on winning seasons. 

{\bf UCLA MAE Lab}\\
 Helped create a script for Python that parses through different news websites to collect the count of specific words while looking at related business articles. These articles are used to perform a sentiment analysis on current economic conditions. Also worked on creating a script that uses the Twitter API to perform similar analysis using tweets. 

\end{rSection}

%----------------------------------------------------------------------------------------
%	TECHNICAL STRENGTHS SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Technical Strengths}

\begin{tabular}{ @{} >{\bfseries}l @{\hspace{4ex}} l }
Software \& Tools: & R, Python, STATA, Tableau, QGIS, SQL, Denodo, LaTeX, Cognos, PowerBI\\
Modeling and Analysis: \ & Asset Pricing, Risk Modeling (Interest, Liquidity, FX, etc...), \\
\ & Econometrics (OLS, ANOVA), Forecasting (ARIMA, VAR, SSM)\\
Technical Knowledge: \ &  Machine Learning (Neural Network, Random Forest, etc), Keras, NLP, \\
\ & Big Data (PCA, Ridge/LASSO, Clustering), ETL processes\\
Languages: \ &  Native: English, Advanced: Spanish, Moderate: French \\
\end{tabular}

\end{rSection}

\end{document}
